Non-analogues in paleoecological reconstruction

### 2015-07-24 Paper_Markdown.Rmd

---
title: "Non-analogues in paleoecological reconstruction - the effects of analogue distance"
author: "Simon Goring *et al*."
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    pandoc_args: "-V geometry:vmargin=1in -V geometry:hmargin=1in"
bibliography: non-analogue_clim.bib
csl: elsevier-harvard.csl
---
Non-analogues in paleoecological reconstruction - the effects of analogue distance
========================================================

@@

Miska Luoto^2,3^

Jack Williams^1,4^

@@

^2^ ?????????

^3^

^4^ Center for Climatic Research, University of Wisconsin-Madison, Madison, WI, USA

@@

Introduction
=========================
Pollen assemblages are widely used for paleoclimatic reconstruction during the Holocene, but this method is also applied to pollen assemblages from the last Glacial period [@bartlein2011pollen, @kaufman2012multi], and earlier [@tarasov2011progress; @brewer2008climate, @cheddadi1998climate].  A particular challenge for climate reconstruction using these much earlier samples is the presence of non-analogue vegetation.  Non-analogue pollen assemblages pose a problem for many reconstruction techniques, these vegetation assemblages may be the result of sets of climate variables that no longer co-exist, ecological conditions resulting from differing rates of species migration (REF) or release from herbivores [@gill2009pleistocene], or as a result of changes in human land use [@goring2015preset].

@@

We have an interesting distribution of loss here.  First, the color coding is based on loss to 'non-analogue' conditions, and we see that it is possible for some samples to have relatively low loss to the no-analogue point, but then much higher losses to the extreme end of our analogue space (see the magenta lines at the very bottom end of Figure 4a.).  Secondly, the points that lose the most samples tend to be in the boreal forest.  To some degree this is surprising, however, the dominance of Pinus and Picea in these forests likely means that they have smaller analogue distances from one another than samples in the southern forests (with more diversity and therfore, likely fewer close analogues), but also, surpisingly, arctic samples also have higer inter-sample distances, and are thus less sensitive to losses of close analogues.
```{r, , echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=5}

>>>>

---
title: "Non-analogues in paleoecological reconstruction"
author: "Simon Goring *et al*."
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    pandoc_args: -V geometry:vmargin=1in -V geometry:hmargin=1in
  md_document:
    variant: markdown_github
  word_document:
    fig_height: 6
    fig_width: 6
csl: elsevier-harvard.csl
pandoc_args: -H margins.sty
bibliography: non-analogue_clim.bib
---
Non-analogues in paleoecological reconstruction - the effects of analogue distance
========================================================

@@

Miska Luoto^2^

Jack Williams^1,3^

@@

^2^ Department of Geosciences and Geography, University of Helsinki, Helsinki, Finland

^3^ Center for Climatic Research, University of Wisconsin-Madison, Madison, WI, USA

@@

Introduction
=========================

Pollen assemblages are widely used for paleoclimatic reconstruction during the Holocene, but this method is also applied to pollen assemblages from the last Glacial period [@bartlein2011pollen, @kaufman2012multi], and earlier [@tarasov2011progress; @brewer2008climate, @cheddadi1998climate].  A particular challenge for climate reconstruction using these much earlier samples is the presence of non-analogue vegetation.  Non-analogue pollen assemblages pose a problem for many reconstruction techniques, these vegetation assemblages may be the result of sets of climate variables that no longer co-exist, ecological conditions resulting from differing rates of species migration (REF) or release from herbivores [@gill2009pleistocene], or as a result of changes in human land use [@goring2015preset].

@@

We have an interesting distribution of loss here.  First, the color coding is based on loss to 'non-analogue' conditions, and we see that it is possible for some samples to have relatively low loss to the no-analogue point, but then much higher losses to the extreme end of our analogue space (see the magenta lines at the very bottom end of Figure 4a.).  Secondly, the points that lose the most samples tend to be in the boreal forest.  To some degree this is surprising, however, the dominance of Pinus and Picea in these forests likely means that they have smaller analogue distances from one another than samples in the southern forests (with more diversity and therfore, likely fewer close analogues), but also, surpisingly, arctic samples also have higer inter-sample distances, and are thus less sensitive to losses of close analogues.

```{r, , echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=5}
